### 2016-09-27 R/ITHIM.Rmd
Changed upper and length in ITHIMObjs chunk
Small bit of cleanup to ITHIM.Rmd
Switched to stratification by State

NHTS.trips <- with(DAYV2PUB,{
    data.frame(HOUSEID = HOUSEID,
                   PERSONID = PERSONID,
                   TDCASEID = TDCASEID,
                   AT = ifelse(TRPTRANS == "22", "cycle",
                        ifelse(TRPTRANS == "23", "walk",
                        ifelse(!is.na(TRPTRANS), "other", NA))),
                   TRVL_MIN = as.numeric(TRVL_MIN),
                   TRVLCMIN = ifelse(as.numeric(TRVLCMIN) != 0, as.numeric(TRVLCMIN), NA),
                   URBRUR  = as.factor(URBRUR),
                   HHSTATE = as.factor(HHSTATE),
                   MSACAT = as.factor(MSACAT),
                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                   TRPTRANS = as.factor(TRPTRANS),
                   TRPMILES = as.numeric(TRPMILES)
)})

@@

```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
#summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
#summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
#summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
#summary(with(subset(NHTS.trips, AT == "walk"), TRVL_MIN))
#summary(with(subset(NHTS.trips, AT == "cycle"), TRVL_MIN))
#summary(with(subset(NHTS.trips, AT == "other"), TRVL_MIN))

walkHours <- with(subset(NHTS.trips, {
    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
}), TRVL_MIN/60)

@@

```{r subset, eval = TRUE, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
```

@@

```{r count, eval = TRUE, echo = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)

@@

```{r CDC, eval = TRUE, echo = FALSE}
wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")

@@

```{r ITHIMCause, eval = TRUE, echo = FALSE}
ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
names(ITHIMcauseKey) <- unique(wonder$ICD10)
wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
```
```{r recodeAge, eval = TRUE, echo = FALSE}
ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
names(ITHIMageKey) <- unique(wonder$WONDERage)
wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
```

@@

```{r wonder, eval = TRUE, echo = FALSE}
wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
wonder <- rbind(wonder, depress)
```

@@

```{r suppressed, eval = TRUE, echo = FALSE, warning = FALSE}
wonder <- within( wonder, {
    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
    })
```

@@

```{r summary2, eval = TRUE, echo = FALSE}
wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
)
```

@@

```{r singleyear, eval = TRUE, echo = FALSE}
pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)

@@

```{r sloppy, eval = TRUE, echo = FALSE}
ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
names(ITHIMageKey2) <- unique(pop14$WONDERage)
pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
```

```{r foop, eval = TRUE, echo = FALSE}
pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))

@@

```{r USGBD, eval = TRUE, echo = FALSE}
USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)

@@

```{r burdenList, eval = TRUE, echo = FALSE}
GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
#saveRDS(GBDList, file = "../data/GBD.rds")
```

@@

```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
upper <- 50
length <- 500

@@

for (HHS in c(paste0("HHS0",1:9),"HHS10")){
    ITHIMList[[HHS]] <- ITHIM.default
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muwt", muwtList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muct", muctList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rwt", RwtList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rct", RctList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
  }

@@

for( muNonTravel in seq(0.01,upper,length.out = length) ){

@@

  for (HHS in c(paste0("HHS0",1:9),"HHS10")){
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", muNonTravel)
  }

@@

barplot(with(NHTS,table(Region)))

@@

ITHIMList <- c(ITHIMList, list("National" = ITHIM.national))

@@

D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
names(D) <- c("ageClass", "mode", "Sex", "time", "Region")

D1 <- subset(D, Region %in% c("HHS01","HHS02","HHS09","HHS10","National") & ageClass != "1" )
D2 <- subset(D, !(Region %in% c("HHS01","HHS02","HHS09","HHS10")) & ageClass != "1" )

p1 <- ggplot(D1, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
p1 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")

p2 <- ggplot(D2, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
p2 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")

>>>>

NHTS.trips <- with(DAYV2PUB,{
    data.frame(HOUSEID = HOUSEID,
                   PERSONID = PERSONID,
                   TDCASEID = TDCASEID,
                   AT = ifelse(TRPTRANS == "22", "cycle",
                        ifelse(TRPTRANS == "23", "walk",
                        ifelse(!is.na(TRPTRANS), "other", NA))),
                   TRVL_MIN = as.numeric(TRVL_MIN),
                   TRVLCMIN = ifelse(as.numeric(TRVLCMIN) != 0, as.numeric(TRVLCMIN), NA),
                   URBRUR  = as.factor(URBRUR),
                   #HHSTATE = as.factor(HHSTATE),
                   MSACAT = as.factor(MSACAT),
                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                   TRPTRANS = as.factor(TRPTRANS),
                   TRPMILES = as.numeric(TRPMILES)
)})

@@

```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
walkHours <- with(subset(NHTS.trips, {
    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
}), TRVL_MIN/60)

@@

```{r subset, eval = TRUE, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
```

@@

```{r count, eval = TRUE, echo = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)

@@

```{r CDC, eval = FALSE, echo = FALSE}
wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")

@@

```{r ITHIMCause, eval = FALSE, echo = FALSE}
ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
names(ITHIMcauseKey) <- unique(wonder$ICD10)
wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
```
```{r recodeAge, eval = FALSE, echo = FALSE}
ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
names(ITHIMageKey) <- unique(wonder$WONDERage)
wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
```

@@

```{r wonder, eval = FALSE, echo = FALSE}
wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
wonder <- rbind(wonder, depress)
```

@@

```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
wonder <- within( wonder, {
    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
    })
```

@@

```{r summary2, eval = FALSE, echo = FALSE}
wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
)
```

@@

```{r singleyear, eval = FALSE, echo = FALSE}
pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)

@@

```{r sloppy, eval = FALSE, echo = FALSE}
ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
names(ITHIMageKey2) <- unique(pop14$WONDERage)
pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
```

```{r foop, eval = FALSE, echo = FALSE}
pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))

@@

```{r USGBD, eval = FALSE, echo = FALSE}
USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)

@@

```{r burdenList, eval = FALSE, echo = FALSE}
GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
#saveRDS(GBDList, file = "../data/GBD.rds")
```

@@

```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
upper <- 100
length <- 100

@@

@@

for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
    ITHIMList[[HHSTATE]] <- ITHIM.default
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
  }

for( muNonTravel in seq(0,upper,length.out = length) ){

@@

  for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
  }

@@

@@

ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))

@@

D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
names(D) <- c("ageClass", "mode", "Sex", "time", "State")

## D1 <- subset(D, Region %in% c("HHS01","HHS02","HHS09","HHS10","National") & ageClass != "1" )
## D2 <- subset(D, !(Region %in% c("HHS01","HHS02","HHS09","HHS10")) & ageClass != "1" )

p <- ggplot(D, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
p + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y")

## p2 <- ggplot(D2, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
## p2 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
